Incorporating periodic variability in hidden Markov models for animal movement

### 2016-02-29 paper3.Rnw
refactor Figure 5 (cat states)

<<state_plots, warning=FALSE, message=FALSE, echo=FALSE, fig.cap="States">>=
temp5 <- tail(getpars(fitsin5s),10)
temp6 <- tail(getpars(fithomo6s),12)
dat1 <- data.frame(log_mean=temp5[c(1,3,5,7,9)],
                     log_sd=temp5[c(2,4,6,8,10)], Model="HMM + TH(sin)") 
dat2 <- data.frame(log_mean=temp6[c(1,3,5,7,9,11)],
                     log_sd=temp6[c(2,4,6,8,10,12)],Model="HMM") 
dat1 <- dat1[order(dat1$log_mean),]
dat2 <- dat2[order(dat2$log_mean),]

dat <- rbind(dat1,dat2)
sd1 <- ggplot(dat,aes(log_mean,log_sd,color=Model))+
    geom_point()+geom_line() + ggtitle("Cat 1")
@

<<state_plots cat2, warning=FALSE, message=FALSE, echo=FALSE>>=
load("cat2sumdat.RData")
temp5 <- tail(getpars(fithomo5s),10)
temp4 <- tail(getpars(fitsin4s),8)
dat1 <- data.frame(log_mean=temp5[c(1,3,5,7,9)],
                     log_sd=temp5[c(2,4,6,8,10)], Model="HMM") 
dat2 <- data.frame(log_mean=temp4[c(1,3,5,7)],
                     log_sd=temp4[c(2,4,6,8)],Model="HMM + TH(sin)") 
dat1 <- dat1[order(dat1$log_mean),]
dat2 <- dat2[order(dat2$log_mean),]

dat <- rbind(dat2,dat1)
sd2 <- ggplot(dat,aes(log_mean,log_sd,color=Model))+
    geom_point()+geom_line() + ggtitle("Cat 2")
@

<<state_plots cat14, warning=FALSE, message=FALSE, echo=FALSE>>=
load("cat14sumdat.RData")
temp5 <- tail(getpars(fithomo5s),10)
temp4 <- tail(getpars(fitsin4s),8)
dat1 <- data.frame(log_mean=temp5[c(1,3,5,7,9)],
                     log_sd=temp5[c(2,4,6,8,10)], Model="HMM") 
dat2 <- data.frame(log_mean=temp4[c(1,3,5,7)],
                     log_sd=temp4[c(2,4,6,8)],Model="HMM + TH(sin)") 
dat1 <- dat1[order(dat1$log_mean),]
dat2 <- dat2[order(dat2$log_mean),]

dat <- rbind(dat2,dat1)
sd14 <- ggplot(dat,aes(log_mean,log_sd,color=Model))+
    geom_point()+geom_line() + ggtitle("Cat 14")
@

<<state_plots cat15, warning=FALSE, message=FALSE, echo=FALSE, fig.cap="State Parameter Estimates">>=
load("cat15sumdat.RData")
temp5 <- tail(getpars(fitsin5s),10)
temp6 <- tail(getpars(fithomo6s),12)
dat1 <- data.frame(log_mean=temp5[c(1,3,5,7,9)],
                     log_sd=temp5[c(2,4,6,8,10)], Model="HMM + TH(sin)") 
dat2 <- data.frame(log_mean=temp6[c(1,3,5,7,9,11)],
                     log_sd=temp6[c(2,4,6,8,10,12)],Model="HMM") 
dat1 <- dat1[order(dat1$log_mean),]
dat2 <- dat2[order(dat2$log_mean),]

dat <- rbind(dat1,dat2)
sd15 <- ggplot(dat,aes(log_mean,log_sd,color=Model))+
    geom_point()+geom_line() + ggtitle("Cat 15")

grid.arrange(sd1,sd2,sd14,sd15)
@

>>>>

<<state_plots, warning=FALSE, message=FALSE, echo=FALSE, fig.cap="Step-length parameters">>=
## translate single fit to data frame of params
trans_pars <- function(fit,npar,Model=NULL) {
    x <- tail(getpars(fit),npar)
    d <- data.frame(log_mean=x[names(x)!="sd"],
                    log_sd=x[names(x)=="sd"])
    if (!is.null(Model)) d$Model <- Model
    d <- arrange(d,log_mean)
    return(d)
}
## get pars for best-fit TH(sin) and HMM models for each cat
get_cat_pars <- function(catnum) {
    load(paste0("cat",catnum,"sumdat.RData"))
    ## figure out
    w <- with(sumdat,which(deltaBIC==min(deltaBIC[model=="HMM + THsin"])))
    optstates <- sumdat$nstates[w]
    ## ASSUME that best fit is het-sin, corresponding homo has +1 states
    sinfit <- get(paste0("fitsin",optstates,"s"))
    homofit <- get(paste0("fithomo",optstates+1,"s"))
    d <- rbind(trans_pars(sinfit,optstates*2,Model="HMM+TH(sin)"),
               trans_pars(homofit,(optstates+1)*2,Model="HMM"))
    return(d)
}
all_cat_pars <- ldply(lme4:::namedList(1,2,14,15),get_cat_pars,.id="catID")
ggplot(all_cat_pars,aes(log_mean,log_sd,color=Model))+
    geom_point()+geom_line() + facet_wrap(~catID,labeller=label_both)
@
